Ms. 2: Consumption and resistance across states and HRRs

### 2017-04-05 states/resistance/ms2.Rmd
moving toward ms2 first linear

I used the Medicare Part D data from 2011. Only beneficiaries greater than 65 years old were used. I also used the antibiogram data. I pooled the antiobiograms from all years.

@@

```{r global_options, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, cache=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE)
import::from(broom, tidy)
library(pander)
library(assertr)
```

@@

@@

```{r load_consumption_data}
state_usage = read_tsv('../bene-consumption/state_usage.tsv') %>%
  filter(year==2011) %>%
  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
  group_by(hrr, drug_group) %>%
  summarize_at(vars(cpkp, did), sum)

hrr_usage = read_tsv('../bene-consumption/hrr_usage.tsv') %>%
  filter(year==2011) %>%
  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
  group_by(hrr, drug_group) %>%
  summarize_at(vars(cpkp, did), sum)

state_dat = left_join(abg, state_usage, by=c('drug_group', 'state'))
hrr_dat = left_join(abg, hrr_usage, by=c('drug_group', 'hrr'))
```

@@

models_result = function(x, this_place) {
  x %>%
    group_by_('bug', 'drug_group', this_place) %>%
    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
              cpkp=head(cpkp, 1),
              did=head(did, 1)) %>%
    group_by(bug, drug_group) %>%
    do(models(., 'mean_nonsus')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    mutate(record_number=1:n(),
           plot_fn=sprintf('plots/%s_%03i.pdf', this_place, record_number),
           plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
    select(bug, drug_group,
           explanatory, correlation=estimate, p.value,
           adj.p.value, sig,
           plot_fn, plot_link)
}

state_res = models_result(state_dat, 'state')
hrr_res = models_result(hrr_dat, 'hrr')

result = bind_rows(
  state_res %>% mutate(place='state'),
  hrr_res %>% mutate(place='hrr')
)
```

@@

#### State-level resistance-averaged correlations

```{r state_average_results_table}
result %>%
  filter(place=='state') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, drug_group,
         explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

@@

```{r state_average_volcano}
result %>%
  filter(place=='state') %>%
  ggplot(aes(x=correlation, y=adj.p.value)) +
  geom_point() +
  geom_hline(yintercept=0.05, col='red') +
  xlab('correlation coefficient') + ylab('p value (reversed)') +
  scale_y_reverse() +
  theme_minimal()
```

#### State-level resistance-averaged correlations using linear models

```{r}
safe_lm = function(data, indep, expl) {
  try(return(lm(sprintf('%s ~ %s', indep, expl), data) %>% tidy), silent=TRUE)
  data_frame(term=expl, estimate=NA, std.error=NA, statistic=NA, p.value=1.0)
}

linear_models = function(df, indep) {
  lapply(c('cpkp', 'did'), function(expl) {
    safe_lm(df, indep, expl) %>%
      filter(term==expl) %>%
      rename(explanatory=term) %>%
      verify(nrow(.) == 1)
  }) %>% bind_rows
}

linear_models_result = function(x, this_place) {
  x %>%
    group_by_('bug', 'drug_group', this_place) %>%
    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
              cpkp=head(cpkp, 1),
              did=head(did, 1)) %>%
    group_by(bug, drug_group) %>%
    do(linear_models(., 'mean_nonsus')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    select(bug, drug_group,
           explanatory, slope=estimate, p.value,
           adj.p.value, sig)
}

linear_state_res = linear_models_result(state_dat, 'state')

linear_state_res %>% kable
```

@@

```{r}
library(mblm)

safe_mblm = function(data, indep, expl) {
  try(return(mblm(sprintf('%s ~ %s', indep, expl), data) %>% tidy), silent=TRUE)
  data_frame(term=expl, estimate=NA, std.error=NA, statistic=NA, p.value=1.0)
}

ts_models = function(df, indep) {
  lapply(c('cpkp', 'did'), function(expl) {
    safe_mblm(df, indep, expl) %>%
      filter(term==expl) %>%
      rename(explanatory=term) %>%
      verify(nrow(.) == 1)
  }) %>% bind_rows
}

ts_models_result = function(x, this_place) {
  x %>%
    group_by_('bug', 'drug_group', this_place) %>%
    summarize(mean_nonsus=mean(percent_nonsusceptible, na.rm=TRUE),
              cpkp=head(cpkp, 1),
              did=head(did, 1)) %>%
    group_by(bug, drug_group) %>%
    do(linear_models(., 'mean_nonsus')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    select(bug, drug_group,
           explanatory, slope=estimate, p.value,
           adj.p.value, sig)
}

ts_state_res = linear_models_result(state_dat, 'state')

ts_state_res %>% kable
```

#### HRR-level resistance-averaged

```{r hrr_average_results_table}
result %>%
  filter(place=='hrr') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, drug_group,
         explanatory, correlation,
         adj.p.value, sig,
         plot_link) %>%
  kable
```

(red line marks $p = 0.05$)

```{r hrr_average_histogram}
result %>%
  filter(place=='hrr') %>%
  ggplot(aes(x=correlation, y=adj.p.value)) +
  geom_point() +
  geom_hline(yintercept=0.05, col='red') +
  xlab('correlation coefficient') + ylab('p value (reversed)') +
  scale_y_reverse() +
  theme_minimal()
```

### All-antibiogram results

In these data, I don't average the resistance before running the correlation, but I do associate each antibiogram with the same geography-wide consumption.

```{r}
all_models_result = function(x, this_place, consumption) {
  x %>%
    mutate_at(vars(cpkp, did), function(x) if_else(is.na(x), 0, x)) %>%
    group_by(bug, drug_group) %>%
    do(models(., 'percent_nonsusceptible')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    select(bug, drug_group,
           explanatory, correlation=estimate, p.value,
           adj.p.value, sig)
}

all_state_res = all_models_result(state_dat, 'state', state_consumption)
all_hrr_res = all_models_result(hrr_dat, 'hrr', hrr_consumption)

all_result = bind_rows(
  all_state_res %>% mutate(place='state'),
  all_hrr_res %>% mutate(place='hrr')
)
```

#### State-wide consumption, all-antibiogram

```{r}
all_result %>%
  filter(place=='state') %>% 
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, drug_group, 
         explanatory, correlation, p.value,
         adj.p.value, sig) %>%
  kable
```

#### State-wide consumtion, all-antibiogram using linear models

```{r}
linear_all_models_result = function(x, this_place) {
  x %>%
    group_by(bug, drug_group) %>%
    do(linear_models(., 'percent_nonsusceptible')) %>%
    ungroup %>%
    filter(!is.na(estimate)) %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value)) %>%
    select(bug, drug_group,
           explanatory, slope=estimate, p.value,
           adj.p.value, sig)
}

linear_all_state_res = linear_all_models_result(state_dat, 'state')
linear_all_hrr_res = linear_all_models_result(hrr_dat, 'hrr')

linear_all_state_res %>% kable
```

#### HRR-wide consumption, all-antibiogram

```{r}
all_result %>%
  filter(place=='hrr') %>% 
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, drug_group, 
         explanatory, correlation, p.value,
         adj.p.value, sig) %>%
  kable
```

>>>>

I used the Medicare Part D data from 2011. Normal filtering was applied to beneficiaries: at least 65 years old, no HMO, 12 months of coverage.

I pooled the antiobiograms from all years.

@@

```{r global_options, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, cache=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE)
import::from(broom, tidy)
import::from(mblm, mblm)
library(pander)
library(assertr)
```

@@

summarize_abg = function(df, place_name) {
  group_by_(df, 'bug', 'drug_group', place_name) %>%
    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
              n_place_antibiograms=n()) %>%
    mutate(place_weight=sqrt(n_place_antibiograms))
}

abg_state_dat = summarize_abg(abg, 'state')
abg_hrr_dat = summarize_abg(abg, 'hrr')

@@

```{r load_consumption_data}
state_usage = read_tsv('../bene-consumption/state_usage.tsv') %>%
  filter(year==2011) %>%
  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
  group_by(state, drug_group) %>%
  summarize_at(vars(cpkp, did), sum) %>%
  ungroup()

hrr_usage = read_tsv('../bene-consumption/hrr_usage.tsv') %>%
  filter(year==2011) %>%
  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
  group_by(hrr, drug_group) %>%
  summarize_at(vars(cpkp, did), sum) %>%
  ungroup()

state_dat = left_join(abg_state_dat, state_usage, by=c('drug_group', 'state')) %>%
  rename(place=state)
hrr_dat = left_join(abg_hrr_dat, hrr_usage, by=c('drug_group', 'hrr')) %>%
  rename(place=hrr)
```

@@

models = function(df, independent_var, explanatory_vars) {
  f = paste0(independent_var, ' ~ ', paste0(explanatory_vars, collapse = ' + ')) %>% formula
  lm(f, data=df) %>% tidy
}

models_result = function(df) {
  group_by(df, bug, drug_group, place) %>%
    summarize(mean_percent_nonsusceptible=weighted.mean(mean_percent_nonsusceptible, place_weight),
              cpkp=unique(cpkp),
              did=unique(did)) %>%
    group_by(bug, drug_group) %>%
    do(models(., 'mean_percent_nonsusceptible', 'cpkp')) %>%
    ungroup %>%
    filter(term=='cpkp') %>%
    mutate(adj.p.value=p.adjust(p.value, method='BH'),
           sig=pvalue_stars(adj.p.value))
   # %>%
    #mutate(record_number=1:n(),
    #       plot_fn=sprintf('plots/%s_%03i.pdf', this_place, record_number),
    #       plot_link=pandoc.link.return(url=plot_fn, text='plot')) %>%
    #select(bug, drug_group,
    #       explanatory, correlation=estimate, p.value,
    #       adj.p.value, sig,
    #       plot_fn, plot_link)
}

state_res = models_result(state_dat)
hrr_res = models_result(hrr_dat)

result = bind_rows(
  state_res %>% mutate(place_type='state'),
  hrr_res %>% mutate(place_type='hrr')
)
```

@@

#### State-level, resistance-averaged linear models

```{r state_average_results_table}
result %>%
  filter(place_type=='state') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, drug_group,
         term, estimate, std.error,
         adj.p.value, sig
         #,plot_link
         ) %>%
  kable
```

@@

```{r state_average_volcano}
result %>%
  filter(place_type=='state') %>%
  ggplot(aes(x=estimate, y=adj.p.value)) +
  geom_point() +
  geom_hline(yintercept=0.05, col='red') +
  xlab('slope') + ylab('p value (reversed)') +
  scale_y_reverse() +
  theme_minimal()
```

#### HRR-level, resistance-averaged linear models

```{r hrr_average_results_table}
result %>%
  filter(place_type=='hrr') %>%
  mutate_if(is.numeric, function(x) signif(x, digits=3)) %>%
  select(bug, drug_group,
         term, estimate, std.error,
         adj.p.value, sig
         #,plot_link
         ) %>%
  kable
```

```{r hrr_average_volcano}
result %>%
  filter(place_type=='hrr') %>%
  ggplot(aes(x=estimate, y=adj.p.value)) +
  geom_point() +
  geom_hline(yintercept=0.05, col='red') +
  xlab('slope') + ylab('p value (reversed)') +
  scale_y_reverse() +
  theme_minimal()
```

@@

swo> Copy the code, and allow a function pointer for lm function to use
